Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-10-27 ms.tex
Fixing some issues w/ the citations, and re-adding Nick's appendix.

We do not include radiative cooling in the entropy equation because
across much of the parameter space radiative cooling is negligible
compared to other sources of cooling (e.g., advection).  Although radiative
cooling cannot be neglected in some cases (especially if $\vwO$ is low
or $\eta$ is high), our 1D scheme cannot account for the true multi-phase behavior of local thermal instabilities anyways (e.g.~\citealt{textit}).  Our steady state solutions neglecting radiation can
nevertheless be used to ascertain over what parameter space cooling
instabilities are likely to develop ($\S\ref{sec:cooling}$).
%Dependence of the heating to cooling ratio has a far steeper
%dependence on vwO than on eta...

@@

Though important across some radii, the importance of conductive heating can be estimated by comparing its magnitude to that of wind heating, viz.~
\begin{eqnarray}
\left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm w}^{2}/2}\right|_{r_{\rm s}} &\sim& \frac{2t_{\rm h}\kappa_0 T^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}(r_{\rm s}) v_{\rm w}^{2}} \nonumber \\
&\sim& 0.5\left(\frac{\eta}{0.1}\right)^{-1}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-\beta} v_{500}^{3}\chi^{3},
\end{eqnarray}
where we have approximated $\nabla^{2} \sim 1/r_{\rm st}^{2}$ and in the second equality have made use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}) and have approximated $4\pi r_{\rm s}^{3}\rho_{\star}(r_{\rm s}) \approx M_{\bullet}(r_{\rm s}/r_{\rm soi})^{\beta}$ with $\beta \approx 1-2$.  We thus see that conductive heating is important near the stagnation radius for high values of $v_{\rm w} \gtrsim 500$ km s$^{-1}$ and $\eta \gtrsim 0.1$, but that it is safely neglected to first order for lower heating rates.

@@

Figure~\ref{fig:vweff} shows an estimate of how the combined heating depends on the age of the stellar population $\tage$ and the SMB mass $\Mbh$.  For $M_{\bullet} \gtrsim 1.5\times 10^8 (\tau_{\star}/t_{\rm h})^{0.5} \Msun$, SNe Ia occur sufficiently frequently within the stagnation radius (i.e., $r_{\rm Ia} > r_{\rm s}$; eq.~[\ref{eq:rIa}]) to contribute their full heating rate.   $v_{\rm w,0}^{\rm Ia} \sim 250$ km s$^{-1}$.  On the other hand, for lower mass BHs Ia SNe are infrequent, in which case main sequence stars ($v_{\rm w,0}^{\star} \simeq 100-150$ km s$^{-1}$) dominate the heating at most times of interest.\footnote{Although SN Ia will still occur surrounding low mass SMBHs, if they are infrequent interior to the radius where the accretion rate is being set, then they will not affect the flow at most near the SMBH at most epochs.}  For very young stellar populations ($\tage \lsim 10^7$ years), $\vwO\simeq 1000$ km s$^{-1}$ due to contributions from line driven winds from high mass main sequence stars.

@@

If cooling becomes important in the flow, it may lead to thermal instabilities.  There are two kinds of ``instabilities", global and local.  The global instability is simply that, if cooling becomes very rapid, the gas loses thermal pressure and starts to inflow faster, i.e. the entire CNM accretes (e.g.~\citealt{Fabian&Peterson06}).  In practice this should could manifest as a large increase in the stagnation radius, due an effective {\it decrease} in the effective heating rate.  This global ``instability" is not a true instability insofar as the gas does not necessarily clump locally into low-T clouds.

@@

\section{Analytic model for dependence of wind heating $v_{\rm w,0}^{\star}$ on stellar age}
\label{app:windheat}
{\bf BDM: Nick add}
  \footnotesize{
    \bibliographystyle{mn2e}
    \bibliography{master}
  }
\end{document}

>>>>

We do not include radiative cooling in the entropy equation because
across much of the parameter space radiative cooling is negligible
compared to other sources of cooling (e.g., advection).  Although
radiative cooling cannot be neglected in some cases (especially if
$\vwO$ is low or $\eta$ is high), our 1D scheme cannot account for the
true multi-phase behavior of local thermal instabilities anyways
(e.g.~\citealt{McCourt+12}).  Our steady state solutions neglecting
radiation can nevertheless be used to ascertain over what parameter
space cooling instabilities are likely to develop
($\S\ref{sec:cooling}$).
%Dependence of the heating to cooling ratio has a far steeper
%dependence on vwO than on eta...

@@

The importance of conductive heating can be estimated by comparing its
magnitude to that of wind heating, viz.~
\begin{eqnarray} \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim& \frac{2t_{\rm h}\kappa_0
T^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}(r_{\rm s}) v_{\rm w}^{2}}
\nonumber \\ &\sim&
0.5\left(\frac{\eta}{0.1}\right)^{-1}\left(\frac{r_{\rm s}}{r_{\rm
soi}}\right)^{-\beta} v_{500}^{3}\chi^{3},
\end{eqnarray}

where we have approximated $\nabla^{2} \sim 1/r_{\rm st}^{2}$ and in the second equality have made use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}) and have approximated $4\pi r_{\rm s}^{3}\rho_{\star}(r_{\rm s}) \approx M_{\bullet}(r_{\rm s}/r_{\rm soi})^{\beta}$ with $\beta \approx 1-2$.  We thus see that conductive heating is important near the stagnation radius for high values of $v_{\rm w} \gtrsim 500$ km s$^{-1}$ and $\eta \gtrsim 0.1$, but that it is safely neglected to first order for lower heating rates.

@@

Figure~\ref{fig:vweff} shows an estimate of how the combined heating depends on the age of the stellar population $\tage$ and the SMB mass $\Mbh$.  For $M_{\bullet} \gtrsim 1.5\times 10^8 (\tau_{\star}/t_{\rm h})^{0.5} \Msun$, SNe Ia occur sufficiently frequently within the stagnation radius (i.e., $r_{\rm Ia} > r_{\rm s}$; eq.~[\ref{eq:rIa}]) to contribute their full heating rate.   $v_{\rm w,0}^{\rm Ia} \sim 250$ km s$^{-1}$.  On the other hand, for lower mass BHs Ia SNe are infrequent, in which case main sequence stars ($v_{\rm w,0}^{\star} \simeq 100-150$ km s$^{-1}$) dominate the heating at most times of interest.\footnote{Although SN Ia will still occur surrounding low mass SMBHs, if they are infrequent interior to the radius where the accretion rate is being set, then they will not affect the flow near the SMBH at most epochs.}  For very young stellar populations ($\tage \lsim 10^7$ years), $\vwO\simeq 1000$ km s$^{-1}$ due to contributions from line driven winds from high mass main sequence stars.

@@

If cooling becomes important in the flow, it may lead to thermal
instabilities.  There are two kinds of ``instabilities", global and
local.  The global instability is simply that, if cooling becomes very
rapid, the gas loses thermal pressure and starts to inflow faster,
i.e. the entire CNM accretes (e.g. ).  In
practice this should could manifest as a large increase in the
stagnation radius, due an effective {\it decrease} in the effective
heating rate.  This global ``instability" is not a true instability
insofar as the gas does not necessarily clump locally into low-T
clouds.

@@

\section{Analytic model for dependence of wind heating $v_{\rm w,0}^{\star}$ on stellar age}
\label{app:windheat}
If we assume that a stellar population forms impulsively in the distant pass with IMF $\mu(m_*)$(with minimum mass $m_0$ and maximum mass $m_1$), then the surviving mass fraction at any future time $t$ is given by 
\begin{equation}
f(t) =\frac{ \int_{m_0}^{m_{\rm max}(t)} m_* \mu(m_*) dm_* }{ \int_{m_0}^{m_1} m_* \mu(m_*) dm_* },
\end{equation}
where 
\begin{equation}
m_{\rm max}(t) \approx 2.5M_\odot~ \left( \frac{t}{10^9~{\rm yr}} \right)^{-0.4}.
\end{equation}
For a Salpeter IMF $\mu(m_*) \propto m_*^{-2.35}$ with $m_0=0.1M_\odot$ and $m_1=100M_\odot$,
\begin{equation}
f_{\rm Sal}(t) = 1.098 - 0.490 \left(\frac{t}{10^{10}~{\rm yr}} \right)^{0.14}
\end{equation}
{\bf NCS: we should probably use a Kroupa/Chabrier IMF, but this gets the ball rolling.}
If we approximate post-main sequence evolution as instantaneous and define $\lambda(m_*)$ as the fractional mass lost during all stages of stellar evolution, then the mass loss rate density
\begin{equation}
q(t) = \frac{\rho_*}{\bar{m}_*} \lambda(m_{\rm max}(t)) m_{\rm max}(t) \frac{df}{dt},
\end{equation}
where the mean stellar mass $\bar{m}_* = \int_{m_0}^{m_{\rm max}(t)} m_*\mu(m_*)dm_* \approx 0.3 M_\odot$.  Further approximating $\lambda(m_*)=0.5$, and using the Salpeter IMF once more, gives
\begin{equation}
q(t) = \frac{\rho_*}{10^{10}~{\rm yr}} \times 0.11 \left(\frac{t}{10^{10}~{\rm yr}} \right)^{-1.26}.
\end{equation}
This is a specific, time-dependent definition of $\eta(t) (=0.11(t/t_{\rm h})^{-1.26})$; if we consider different star formation scenarios (for example, continuous star formation) or different IMFs, it will change.  Once these free parameters are specified, however, we can answer an important question: do young stellar populations increase or decrease the SMBH feeding rate $\dot{M}$?  Clearly, $\eta(t)$ is larger for young stellar populations, but these stars also have high wind velocities that diminish the stagnation radius.  Crudely approximating $v_{\rm w}=200~{\rm km~s}$ for $m_{\rm max} < 10M_{\odot}$ and $v_{\rm w}=1000~{\rm km~s}$ for $m_{\rm max} > 10M_{\odot}$ (motivated by the transition from dust-driven wind loss on the AGB to line-driven wind loss on the main sequence), we can employ the relation $\dot{M} \propto \eta(t) r_{\rm s}^{2-\Gamma}\propto \eta(t) v_{\rm w}^{-4+2\Gamma}$ (where $\rho_* \propto r^{-\Gamma}$) to determine the impact of stellar ``youth'' on SMBH feeding rates.

\begin{figure}
\includegraphics[width=\columnwidth]{NickPlot.eps}
\caption{\label{NickPlot} SMBH feeding rates $\dot{M}=\eta(t) \times M_*(r_{\rm s})$, in arbitrary units.  The green, blue, and black curves are for galaxies with $\Gamma$ values of $0.1$, $0.5$, and $0.9$, respectively.  Solid curves represent impulsive-mode star formation, while dotted curves represent continuous-mode star formation.}
\end{figure}

In Fig. \ref{NickPlot} we plot $\dot{M}$, in arbitrary units, as a function of time, for three different stellar density profiles $\Gamma = \{1.1, 1.5, 1.9\}$ (which are normalized to have the same mass at an influence radius $r_{\rm soi}=10~{\rm pc}$ around a $10^7M_\odot$ SMBH).  We parametrize the wind velocity as 
\begin{equation}
\frac{v_{\rm w}}{\rm km~s^{-1}}=600-400\tanh\left( \frac{t-10^{7.5}~{\rm yr}}{10^{7}~{\rm yr}}\right).
\end{equation}
The ``impulsive burst'' mode of star formation produces large ($\sim 10$) differences between the three $\dot{M}$ curves at early times, when $r_{\rm s}$ is small, but smaller ($\sim 3$) differences at late times, when $r_{\rm s}$ is large.  We also plot, as dotted curves, a simple model for the ``continuous'' mode of star formation, where mass loss is calculated as $\bar{\eta} = \int\eta(t)dt/t_{\rm h} \approx 4$ and an average energy injection in the wind is calculated as $\bar{v_{\rm w}^2}=\int\eta(t)v_{\rm w}^2(t)dt/\bar{\eta} \approx (800~{\rm km~s}^{-1})^2$.  Interestingly, the continuous mode of star formation produces small differences from late-time $\dot{M}$ seen in cusp galaxies with impulsive star formation; however, continuous mode star formation decreases late-time $\dot{M}$ by an order of magnitude relative to impulsive star formation in core galaxies.
  \footnotesize{
    \bibliographystyle{mn2e}
    \bibliography{master}
  }
\end{document}
